---
title: "HW1"
author: "Audrey Bahr"
date: "March 10, 2021"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r}
library(tidyverse)
```

**Challenge 1**

This challenge involves string manipulation on the last chapter of Charles Darwin's "On the Origin of Species."
```{r}
darwin <- read.delim("darwin.txt", comment.char="#", stringsAsFactors=FALSE)
```

```{r}
darwin_vec <- darwin$CHAPTER.XV..RECAPITULATION.AND.CONCLUSION.

# print number of paragraphs
print(length(darwin_vec))

# print 34th paragraph
print(darwin_vec[34])

# split vector of paragraphs into a vector of words
darwin_split <- strsplit(darwin_vec, " ") %>% unlist()
# remove punctuation
darwin_splitpunct <- gsub("[[:punct:]]","", darwin_split)

# count number of different words
length(unique(darwin_splitpunct))

# find most common word
row <- which.max(data.frame(table(darwin_splitpunct))$Freq)
data.frame(table(darwin_splitpunct))[row,]

# words that appear a certain number of times
sum(data.frame(table(darwin_splitpunct))$Freq == 1)
sum(data.frame(table(darwin_splitpunct))$Freq >= 5) 

# last paragraph
final_quote <- darwin_vec[56]
split_quote <- strsplit(final_quote, " ") %>% unlist()

# extract every third word
every_third <- split_quote[seq(3, length(split_quote), by = 3)]
print(every_third)

# sort by reverse alphabetical order
every_third <- gsub("[[:punct:]]","", every_third)
every_third <- sort(every_third, TRUE)

print(every_third)
```

There are 56 paragraphs and 2,058 unique words in this vector. "The" is the most common word in this chapter, appearing 718 times. There are 1071 words that appear only once, and 369 words that appear five times or more. Punctuation was removed when the frequency of words were counted, because otherwise words with punctuation attached are counted as different (e.g. condition != condition,).

**Challenge 2**

Average high temperatures in January for six cities.
```{r}
t <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

names(t) <- city

t[1:3]
t[c("Paris", "San Juan")]
```

**Challenge 3**

Matrix manipulation.

```{r}
m1 <- matrix(data = sort(c(0:159), TRUE), nrow = 8, ncol = 20)
print(m1[5,2])
print(m1[5:7,])

m2 <- m1[3:6,4:9]
class(m2)
mode(m2)
```
The new object, m2, is a matrix of mode numeric.

**Challenge 4**

Array manipulation.

```{r}
a <- array(sort(seq(2, 800, 2), TRUE), dim = c(5, 5, 4, 4))

a[1, 1, 1, 2]
a[2, 3, 2, ]
a[1:5, 1:5, 3, 3]
```

**Challenge 5**

Simplified primate taxonomy represented by a multidimensional list.

```{r}
# superfamilies
Lorisoidea <- c("Lorisidae", "Galagidae")
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae", "Daubentoniidae")
Tarsioidea <- c("Tarsiidae")
Ceboidea <- c("Cebidae", "Atelidae", "Pitheciidae")
Hominoidea <- c("Hylobatidae", "Hominidae")
Cercopithecoidea <- c("Cercopithecidae")

# parvorders
Catarrhini <- list("Superfamily Hominoidea" = Hominoidea, "Superfamily Cercopithecoidea" = Cercopithecoidea)
Platyrrhini = list("Superfamily Ceboidea" = c(Ceboidea))

# infraorders
Lorisiformes = list("Superfamily Lorisoidea" = (Lorisoidea))
Lemuriformes = list("Superfamily Lemuroidea" = (Lemuroidea))
Tarsiiformes = list("Superfamily Tarsioidea" = (Tarsioidea))
Simiiformes = list("Parvorder: Platyrrhini" = Platyrrhini, "Parvorder: Catarrhini" = Catarrhini)

# suborders
Strepsirhini = list("Infraorder: Lorisiformes" = Lorisiformes, "Infraorder: Lemuriformes" = Lemuriformes)
Haplorhini = list("Infraorder: Tarsiiformes" = (Tarsiiformes), "Infraorder: Simiiformes" = Simiiformes)

# all together
Primates = list("Suborder: Strepsirhini" = Strepsirhini, "Suborder: Haplorhini" = Haplorhini)

# extract elements
platyrrhines <- Primates[[2]][[2]][1]
class(platyrrhines)
mode(platyrrhines)

Primates$`Suborder: Haplorhini`$`Infraorder: Tarsiiformes`$`Superfamily Tarsioidea`
```

The class and mode of platyrrhines are both list.
